Actuator line modeling of vertical-axis turbines

### 2016-05-13 paper.tex
Move comments about shaft foil data up
Remove sinusoidal tsr comments and save computational expense for the conclusions

\section{Static foil coefficient data}

@@

@@

The same foil coefficient data were used for all simulations---those for a NACA
0021 as reported by Sheldahl and Klimas \cite{Sheldahl1981}. Each rotor's shaft
was assumed to have a drag coefficient $C_d = 1.1$, and the blade support strut
end element drag coefficients were set to 0.05, to approximate the effects of
separation in the corners of the blade--strut connections.

Since the ALM is intended to be an engineering tool when coupled with RANS, it
was assumed that information about tip speed ratio due to control details---that
is, sinusoidally oscillating $\lambda$---would not be known a priori, and was
excluded, unlike the 3-D blade-resolved cases reported in
\cite{Bachant2016-BR-CFD}. Note that a systematic investigation of the effects
of sinusoidal $\lambda$ was not undertaken, but for the UNH-RVAT a one to two
percentage point increase in $C_P$ was observed when running with similar
parameters as the blade-resolved simulation.

Compared to 3-D blade-resolved RANS, the ALM can solve a standalone turbine case
on the order of 0.1 CPU hours per second of simulated time versus 1,000 CPU
hours per second---a savings of 4 orders of magnitude.

@@

An actuator line model for cross-flow turbines, including a Leishman--Beddoes
type dynamic stall model, flow curvature, added mass, and lifting-line based end
effects corrections, was developed and validated against experimental datasets
acquired for high and medium solidity rotors at scales where the performance and
near-wake dynamics were essentially Reynolds number independent. When coupled to
a $k$--$\epsilon$ RANS solver ALM simulations took $O(0.1)$ CPU hours per second
of simulated time, while when coupled with a Smagorinsky LES model the computing
time was $O(10)$ hours per second, which represent a four and two order of
magnitude decrease in computational expense versus 3-D blade-resolved RANS,
respectively.

>>>>

\section{Static profile coefficient data}

@@

Each rotor's shaft was assumed to have a drag coefficient $C_d = 1.1$, and the
blade support strut end element drag coefficients were set to 0.05, to
approximate the effects of separation in the corners of the blade--strut
connections.

@@

@@

An actuator line model for cross-flow turbines, including a Leishman--Beddoes
type dynamic stall model, flow curvature, added mass, and lifting-line based end
effects corrections, was developed and validated against experimental datasets
acquired for high and medium solidity rotors at scales where the performance and
near-wake dynamics were essentially Reynolds number independent. When coupled to
a $k$--$\epsilon$ RANS solver ALM simulations took $O(0.1)$ CPU hours per second
of simulated time, while when coupled with a Smagorinsky LES model the computing
time was $O(10)$ hours per second, which represent a four and two order of
magnitude decrease in computational expense versus 3-D blade-resolved
RANS~\cite{Bachant2016-BR-CFD}, respectively.
